---
title: "GapMinderData"
author: "Josh Levy"
date: "3/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
```

```{r}
gapminder <- read_csv("https://raw.githubusercontent.com/jules32/2017-11-30-MBARI/gh-pages/data/gapminder.csv")
```

Show first 2 and bottom 7 rows of data table
```{r}
head(gapminder, 2)
tail(gapminder,7)
```

```{r}
str(gapminder)
```

```{r}
dim(gapminder)
ncol(gapminder)
nrow(gapminder)
```

```{r}
unique(gapminder$country)
```

```{r}
summary(gapminder)
```

```{r}
# install.packages("skimr")
library("skimr", lib.loc="~/Library/R/3.2/library")
skim(gapminder)
```

filter by certain variable metrics
```{r}
filter(gapminder, lifeExp < 30, gdpPercap < 4000)
```

select columns of interest. If function is assigned as new variable, continent will be removed and not able to be re added. 
```{r}
select(gapminder, country, year, pop)
```

remove columns and filter using same dataset. Inject result of first step into second step
```{r}

select(gapminder, -continent) %>% 
  filter(lifeExp < 30, gdpPercap < 4000)
```

create new column based of operation using two already existing columns and arrange columns based on new column from hi to low
```{r}
gapminder %>%
  mutate(gdp = gdpPercap * pop) %>%
  arrange(desc(gdp))
```

